---
title: "Homework 3"
author: "Julia Thompson"
date: "10/14/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(ggridges)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Load the data:

```{r}
library(p8105.datasets)
data("instacart")
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

```{r}
# Look at a random person's order info for an illustrative example
set.seed(5)
instacart_sample = instacart %>% 
  filter(user_id == sample(pull(instacart, user_id), 1)) 
```

The instacart dataset contains detailed information from online grocery orders from 2017. It contains `r nrow(instacart)` observations and `r ncol(instacart)` variables. The dataset is organized by order. Within each order, there is one row for each item purchased, and these items have  an ID and description given by _product_id_ and _product_name_. Each customer and order have a unique ID, represented by the variables _user_id_ and _order_id_. The dataset also contai
ns IDs and names for each aisle and department from which every item came. There are `r n_distinct(pull(instacart, order_id))` distinct orders, made by `r n_distinct(pull(instacart, user_id))` different customers. Therefore, this dataset contains one order per customer. There is not a case of the same customer havind multiple orders here. 

For example, we can look at the order for the customer with _user_id_ `r min(pull(instacart_sample, user_id))`. That customer ordered `r nrow(instacart_sample)` items, including "`r min(pull(instacart_sample, product_name))`." Customer `r min(pull(instacart_sample, user_id))` had ordered `r sum(pull(instacart_sample, reordered))` of the `r nrow(instacart_sample)` items previously. 

How many aisles are there, and which aisles are the most items ordered from?

```{r}
# get number of unique aisles
n_distinct(pull(instacart, aisle_id))

# sort this, and then show the top 3
top_aisles = 
  instacart %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>% 
  top_n(3, n)

knitr::kable(top_aisles)
```

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
aisles_over_10k = 
  instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  arrange(desc(n))

ggplot(aisles_over_10k, aes(x = reorder(aisle, -n), y = n)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    title = "Number of Items Ordered in Each Aisle",
    x = "Aisle",
    y = "Number of Items",
    caption = "Number of items ordered in each aisle with more than 10,000 items ordered"
  )
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
fav_items = instacart %>% 
  filter(aisle == c("baking ingredients","dog food care","packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(
    n = n()
  ) %>% 
  top_n(3, n) %>% 
  arrange(aisle, n)

knitr::kable(fav_items)
```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
hr_of_day = instacart %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_hr = mean(order_hour_of_day)
  ) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hr
  ) %>% 
  rename(
    Sunday    = `0`, 
    Monday    = `1`, 
    Tuesday   = `2`, 
    Wednesday = `3`, 
    Thursday  = `4`, 
    Friday    = `5`, 
    Saturday  = `6`
  )

knitr::kable(hr_of_day)
```

