---
title: "Homework 3"
author: "Julia Thompson"
date: "10/14/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Load the data:

```{r}
library(p8105.datasets)
data("instacart")
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

```{r}
# Look at a random person's order info for an illustrative example
set.seed(5)
instacart_sample = instacart %>% 
  filter(user_id == sample(pull(instacart, user_id), 1)) 
```

The instacart dataset contains detailed information from online grocery orders from 2017. It contains `r nrow(instacart)` observations and `r ncol(instacart)` variables. The dataset is organized by order. Within each order, there is one row for each item purchased, and these items have  an ID and description given by _product_id_ and _product_name_. Each customer and order have a unique ID, represented by the variables _user_id_ and _order_id_. The dataset also contai
ns IDs and names for each aisle and department from which every item came. There are `r n_distinct(pull(instacart, order_id))` distinct orders, made by `r n_distinct(pull(instacart, user_id))` different customers. Therefore, this dataset contains one order per customer. There is not a case of the same customer havind multiple orders here. 

For example, we can look at the order for the customer with _user_id_ `r min(pull(instacart_sample, user_id))`. That customer ordered `r nrow(instacart_sample)` items, including "`r min(pull(instacart_sample, product_name))`." Customer `r min(pull(instacart_sample, user_id))` had ordered `r sum(pull(instacart_sample, reordered))` of the `r nrow(instacart_sample)` items previously. 
